# Data transformation

```{r}
library("ggplot2")
library("dplyr")
library("DT")
library("xtable")
library("knitr")

```


## Extraction

The datasets that we have downloaded from their respective sites have a lot of columns, many of which are not needed for this project.

```{r}
Fao<- read.csv("./datasets/FAOSTAT_data_4-10-2021.csv")
cat(colnames(Fao),sep="\n",append=FALSE)
```

The columns for the Forest Land data look like this:

||||||
|----|----|----|----|----|----|----|
ï..Domain.Code |Domain |Area.Code |Area |Element.Code |Element |Item.Code|
chr |chr |int |chr |int |chr |int|

||||||
|----|----|----|----|----|----|----|
Item |Year.Code |Year |Unit |Value |Flag |Flag.Description
chr |int |int |chr |num |chr |chr|

## Transformation

On randomly plotting a scatter plot of forest area for the top countries by area. That is 
> \ 
1. Russia \
2. Canada \
3. China \
4. US \
5. Brazil \
6. Australia \
7. India \
8. Argentina \
9. Kazakhstan \
10. Algeria \

We get the following plots.


```{r,null_prefix=TRUE}
kable(Fao[Fao$Area=="Russian Federation",][,c("Area","Year","Value")])
plot(Value ~ Year , data=Fao[Fao$Area=="Russian Federation",])
kable(Fao[Fao$Area=="Canada",][,c("Area","Year","Value")])
plot(Value ~ Year , data=Fao[Fao$Area=="Canada",])
kable(Fao[Fao$Area=="China",][,c("Area","Year","Value")])
plot(Value ~ Year , data=Fao[Fao$Area=="China",])
kable(Fao[Fao$Area=="United States of America",][,c("Area","Year","Value")])
plot(Value ~ Year , data=Fao[Fao$Area=="United States of America",])
kable(Fao[Fao$Area=="Brazil",][,c("Area","Year","Value")])
plot(Value ~ Year , data=Fao[Fao$Area=="Brazil",])
kable(Fao[Fao$Area=="Australia",][,c("Area","Year","Value")])
plot(Value ~ Year , data=Fao[Fao$Area=="Australia",])
kable(Fao[Fao$Area=="India",][,c("Area","Year","Value")])
plot(Value ~ Year , data=Fao[Fao$Area=="India",])
kable(Fao[Fao$Area=="Argentina",][,c("Area","Year","Value")])
plot(Value ~ Year , data=Fao[Fao$Area=="Argentina",])
kable(Fao[Fao$Area=="Kazakhstan",][,c("Area","Year","Value")])
plot(Value ~ Year , data=Fao[Fao$Area=="Kazakhstan",])
kable(Fao[Fao$Area=="Algeria",][,c("Area","Year","Value")])
plot(Value ~ Year , data=Fao[Fao$Area=="Algeria",])
```

```{r}

```


## Loading