# Missing values

```{r,null_prefix=TRUE}

library(dplyr, warn.conflicts = FALSE)
library(ggplot2)
library(tidyverse)
library(knitr)
Fao<- read.csv("./datasets/FAOSTAT_data_4-10-2021.csv")
Fao<- Fao[,c("Area","Year","Value")]
Fao<-rename(Fao, "Country" = "Area")
```

First lets check which of the three columns have missing values.

```{r,null_prefix=TRUE}
cat(sapply(Fao, function(x) sum(is.na(x))),sep="\n")
```

Thus the third column that is Value has 537 missing values.
Now lets see which countries have these.

```{r}
missing <- Fao[is.na(Fao$Value),]
cat(unlist(unique(missing[c("Country")])),fill = getOption("width"),sep=", ")
```

```{r}
missing
ggplot(missing, aes(x=Country, fill = Value, group = Value)) 
```


Now comes the historical context to these missing values.

>1. Countries such as Armenia, Azerbaijan came into existence in 1991 so have data missing for 1991. \
2. Similarly the Soviet Union broke into 15 sovereign states. All 15 of which are in the above list. \
3. Other countries also follow the same logic. South Sudan came into existence in 2011. For the time being we shall delete these countries as they don't affect our comparisons much. \ 
4. Even countries which broke into two separate countries have missing values. \